---
title: "anric"
format: html
editor_options: 
  chunk_output_type: console
---
```{r}

patients |> summarise(across(everything(), ~ sum(is.na(.)))) |>
  pivot_longer(everything(), names_to = "column", values_to = "count") |>
  arrange(desc(count)) |>
  kable()
```

```{r}

diagnosis |> summarise(across(everything(), ~ sum(is.na(.)))) |>
  pivot_longer(everything(), names_to = "column", values_to = "count") |>
  arrange(desc(count)) |>
  kable()
```

```{r patients-cols}
names(patients)
```

```{r patients-vis}
ggplot(patients, aes(x = first_race, fill = first_race)) +
  geom_bar() +
  scale_y_log10() 

ggplot(patients, aes(x = "", fill = first_race)) +
  geom_bar(position = "fill")
ggplot(patients, aes(x = "", fill = vital_status)) +
  geom_bar(position = "fill")

```

```{r}
block_group <- patients |>
  count(census_block_group_fips_code)
```
```{r}
unique(patients$my_chart_status)
patients |>
  filter(is.na(patients$my_chart_status))
```


```{r}
head(encounters, n = 10)
```


```{r}
encounters_patients <- left_join(encounters, patients,
  by = join_by(patient_durable_key == durable_key))
```


```{r}
virtual <- encounters_patients |>
  filter(visit_type_description == "Clinic Visits- Virtual" 
         | visit_type_description == "Virtual Visits"
    | visit_type_description == "Clinic Visit- Virtual") 

virtual_patients <- virtual |>
  select(
    patient_durable_key, census_block_group_fips_code:patient_birth_year_bin) |>
  distinct()

```

```{r}
virtual_patients |>
  filter(is.na(census_block_group_fips_code)) |>
  distinct()
```


```{r}
virtual_diagnosis <- virtual |> 
  left_join(diagnosis, by = join_by(primary_diagnosis_key == diagnosis_key))
```


# patient demographics:

Racial: The vast majority of virtual appointments are made by White patients. 
There is also a much smaller proportion of NAs in the virtual data, and the 
proprtions of each racial minority relative to each other are relativly similar
between the two datasets. When visualized, it almost looks like the vast majority
of the NA's were 'converted' to 
```{r}
race_patients <- patients |>
  group_by(first_race) |>
  summarise(
    count = n(),
    proportion = n() / nrow(patients),
    .groups = "drop"
  )
race_patients <- virtual_patients |>
  group_by(first_race) |>
  summarise(
    count = n(),
    proportion = n() / nrow(virtual_patients),
    .groups = "drop"
  ) |>
  rename(
    virtual_count = count,
    virtual_proportion = proportion
  ) |>
  left_join(race_patients, by = join_by(first_race))
```

```{r patients-vis}
ggplot(virtual_patients, aes(x = first_race, fill = first_race)) +
  geom_bar()

ggplot(patients, aes(x = first_race, fill = first_race)) +
  geom_bar() 
ggplot(virtual_patients, aes(x = "", fill = first_race)) +
  geom_bar(position = "fill")

ggplot(patients, aes(x = "", fill = first_race)) +
  geom_bar(position = "fill")

```

Legacy code:
```{r}

dead_encounters |> summarise(across(everything(), ~ sum(is.na(.)))) |>
  pivot_longer(everything(), names_to = "column", values_to = "count") |>
  arrange(desc(count)) |>
  kable()

```



```{r}
diagnosis |> count(group_name)
diagnosis |> count(diagnosis_key) |> arrange(desc(n))

# Step 1 & 2: Count distinct B per A
diagnosis_conflict <- diagnosis |>
  group_by(diagnosis_key) |>
  summarise(
    n_B = n_distinct(group_name),          # number of unique B values
    B_values = list(unique(group_name)),   # actual B values
    .groups = "drop"
  ) |>
  filter(n_B > 1)

# View result
diagnosis_conflict
```


```{r}
#dead_encounters <- inner_join(encounters, dead_patients, by = join_by(patient_durable_key == durable_key)) |>
  inner_join(diagnosis, join_by(diagnosis))

```



```{r}
dead_patients <- patients |>
  filter(vital_status == "Deceased")
unique(dead_patients$patient_birth_year_bin)

ggplot(dead_patients, aes(x = patient_birth_year_bin)) +
  geom_bar() +
  scale_y_log10()
```
