---
title: "anric"
format: html
editor_options: 
  chunk_output_type: console
---
```{r}

patients |> summarise(across(everything(), ~ sum(is.na(.)))) |>
  pivot_longer(everything(), names_to = "column", values_to = "count") |>
  arrange(desc(count)) |>
  kable()
```

```{r}

diagnosis |> summarise(across(everything(), ~ sum(is.na(.)))) |>
  pivot_longer(everything(), names_to = "column", values_to = "count") |>
  arrange(desc(count)) |>
  kable()
```

```{r patients-cols}
names(patients)
```

```{r patients-vis}
ggplot(patients, aes(x = first_race, fill = first_race)) +
  geom_bar() +
  scale_y_log10() 

ggplot(patients, aes(x = "", fill = first_race)) +
  geom_bar(position = "fill")
ggplot(patients, aes(x = "", fill = vital_status)) +
  geom_bar(position = "fill")

```

```{r}
block_group <- patients |>
  count(census_block_group_fips_code)
```



```{r}
head(encounters, n = 10)
```


```{r}
encounters_patients <- left_join(encounters, patients,
  by = join_by(patient_durable_key == durable_key))
```


```{r}
virtual <- encounters_patients |>
  filter(visit_type_description == "Clinic Visits- Virtual" 
         | visit_type_description == "Virtual Visits"
    | visit_type_description == "Clinic Visit- Virtual") 

virtual_patients <- virtual |>
  select(
    patient_durable_key, first_race:patient_birth_year_bin) |>
  distinct()

```



```{r}
virtual_patients |>
  filter(is.na(census_block_group_fips_code)) |>
  distinct()
```


```{r}
virtual_diagnosis <- semi_join(diagnosis, virtual, by = c("diagnosis_key" = "primary_diagnosis_key"))
```

```{r}

virtual |> summarise(across(everything(), ~ sum(is.na(.)))) |>
  pivot_longer(everything(), names_to = "column", values_to = "count") |>
  arrange(desc(count)) |>
  kable()
```

```{r}
diagnosis_conflict <- virtual_diagnosis |>
  group_by(diagnosis_key) |>
  summarise(
    n_B = n_distinct(group_name),          # number of unique B values
    B_values = list(unique(group_name)),   # actual B values
    .groups = "drop"
  ) |>
  filter(n_B > 1)

# View result

```


# patient demographics:

Racial: The vast majority of virtual appointments are made by White patients. 
There is also a much smaller proportion of NAs in the virtual data, and the 
proprtions of each racial minority relative to each other are relativly similar
between the two datasets. When visualized, it almost looks like the vast majority
of the NA's were 'converted' to 
```{r}
race_patients <- patients |>
  group_by(first_race) |>
  summarise(
    count = n(),
    proportion = n() / nrow(patients),
    .groups = "drop"
  )
race_patients <- virtual_patients |>
  group_by(first_race) |>
  summarise(
    count = n(),
    proportion = n() / nrow(virtual_patients),
    .groups = "drop"
  ) |>
  rename(
    virtual_count = count,
    virtual_proportion = proportion
  ) |>
  left_join(race_patients, by = join_by(first_race))
```

```{r patients-vis}
ggplot(virtual_patients, aes(x = first_race, fill = first_race)) +
  geom_bar()

ggplot(patients, aes(x = first_race, fill = first_race)) +
  geom_bar() 
ggplot(virtual_patients, aes(x = "", fill = first_race)) +
  geom_bar(position = "fill")

ggplot(patients, aes(x = "", fill = first_race)) +
  geom_bar(position = "fill")

```


group names:

```{r}
# isolate diagnosis keys from virtual
# isolate diagnosis keys and group names from diagnosis
# join

virtual_slice <- virtual |> select(primary_diagnosis_key)
virtual_diag_slice <- virtual_diagnosis |> select(diagnosis_key, group_name) |> distinct()

virtual_diagnosis_stats <- left_join(virtual_slice, virtual_diag_slice, 
  by = c("primary_diagnosis_key" = "diagnosis_key"),
  relationship = "many-to-many", multiple = "all")


virtual_group_name_stat <- count(virtual_diagnosis_stats, group_name) |>
  rename(virtual_count = n)



```


# diagnosis
some diagnosis groups that are most commonnly virtual include general visits 
(no issues), eating disorders, obesity / weight management, and pregnancy management.
Some diagnosis gorups that are almost never virtual include NA,
(arthritis, bone fractures, mamograams, labor, wounds, etc). 
```{r}
# isolate diagnosis keys from virtual
# isolate diagnosis keys and group names from diagnosis
# join

encounter_slice <- encounters |> select(primary_diagnosis_key)
diag_slice <- diagnosis |> select(diagnosis_key, group_name) |> distinct()

diagnosis_stats <- left_join(encounter_slice, diag_slice, 
  by = c("primary_diagnosis_key" = "diagnosis_key"),
  relationship = "many-to-many", multiple = "all")


group_name_stat <- count(diagnosis_stats, group_name) |>
  left_join(virtual_group_name_stat, by = join_by(group_name)) |>
  rename(virtual_count = n.y, count = n.x) |>
  mutate(proportion = virtual_count * 100 / count)


```

# age
older patients are more likely to have virtual appointments?

```{r}
virtual_patients |>
  semi_join(patients, by = join_by(patient_durable_key = durable_key)
```

```{r patients-vis}


ggplot(virtual_patients, aes(x = patient_birth_year_bin)) +
  geom_histogram() + 
  labs(title = "virtual patients")

ggplot(patients, aes(x = patient_birth_year_bin)) +
  geom_histogram() +
    labs(title = "All patients")


```


```{r}

combined_data <- bind_rows(
  virtual_patients %>% mutate(group = "virtual_patients"),
  patients        %>% mutate(group = "all_patients")
)

age_summary <- combined_data %>%
  group_by(group, patient_birth_year_bin) %>%
  summarise(count = n(), .groups = "drop") %>%
  arrange(group, patient_birth_year_bin)
age_summary <- age_summary %>%
  pivot_wider(
    names_from  = group,
    values_from = count,
    values_fill = 0
  ) |>
  mutate(prop = virtual_patients / all_patients )
```



# Legacy code:
```{r}

dead_encounters |> summarise(across(everything(), ~ sum(is.na(.)))) |>
  pivot_longer(everything(), names_to = "column", values_to = "count") |>
  arrange(desc(count)) |>
  kable()

```



```{r}
#dead_encounters <- inner_join(encounters, dead_patients, by = join_by(patient_durable_key == durable_key)) |>
  inner_join(diagnosis, join_by(diagnosis))

```



```{r}
dead_patients <- patients |>
  filter(vital_status == "Deceased")
unique(dead_patients$patient_birth_year_bin)

ggplot(dead_patients, aes(x = patient_birth_year_bin)) +
  geom_bar() +
  scale_y_log10()
```


```{r}
# Add a 'group' column to each data frame, then combine
virtual_patients$group <- "Virtual patients"
patients$group <- "In person patients"

combined <- rbind(
  virtual_patients[, c("patient_birth_year_bin", "group")],
  patients[, c("patient_birth_year_bin", "group")]
)


```

```{r}
ggplot(combined, aes(x = patient_birth_year_bin, fill = group)) +
  geom_histogram(position = "identity") +
  scale_fill_manual(values = c("In person patients" = "steelblue", "Virtual patients" = "coral")) +
  labs(
    title = "Patients by encounter type",
    x = "Patient birth year bin",
    y = "Count (square rooted)",
    fill = NULL
  ) +  scale_y_continuous(trans = "sqrt") +  theme_minimal() +
  theme(legend.position = "top")

```

